---
title: "Machine learning based on Random forest"
output: html_document
---

## Introduction


Feature selection using Random forest with Boruta on the DEGs.


```{r setup, include=TRUE, warning = F, message = F}
knitr::opts_chunk$set(warning = F, message = F)
library(tidyverse)
library(Biobase)
library(randomForest)
library(ggplot2)
library(data.table)
library(caret)
library(mlbench)
library(Boruta)
library(pROC)

library(sysfonts)
library(showtextdb)
library(showtext)
#激活显示中文的包
showtext_auto()

rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 10000 * 1024^2)

grp_names <- c("Healthy", "PAAD")
grp_colors <- c("#1F78B4", "#EE2B2B")
fig_plot <- "../../Result/JiaBo/figure/"
```


## Importing data 
```{r}
load("../../Result/Exp/exo.RData")
load("../../Result/JiaBo/2DEG/overlap_genes.rda")
source("./utilities.R")
source("../functions.R")

item = "exo" 
profile <- exp_dat_exo
metadata <- phen_dat_exo  
metadata <- metadata %>% dplyr::rename(Group = Compvar)
pref <- paste0("../../Result/JiaBo/3ML/", item)
fig_plot <- "../../Result/JiaBo/figure/ML1/"
pref_plot <- paste0(fig_plot, item, "_RF_")

profile_DEG <- profile[pmatch(overlap_genes, rownames(profile)), ] %>%
  t() %>%
  as.data.frame()
```

+ Remove redundant variables
```{r}
feature_cor <- Hmisc::rcorr(as.matrix(profile_DEG), type = "spearman")
index_cor <- caret::findCorrelation(feature_cor$r, cutoff = 0.75, verbose = TRUE)
feature_remain <- feature_cor$r[-index_cor, -index_cor]
profile_remain <- profile_DEG %>% 
  dplyr::select(all_of(rownames(feature_remain))) 

print(dim(profile_remain))

write.table(profile_remain, paste0(pref, "_RF_RM_profile.tsv"),
            row.names = T, sep = "\t", quote = F)
```

+ MergeData 

```{r}
MergeData <- metadata %>%
  dplyr::select(SampleID, Group) %>%
  dplyr::inner_join(profile_remain %>%
                      tibble::rownames_to_column("SampleID"),
                    by = "SampleID") %>%
  tibble::column_to_rownames("SampleID") %>%
  dplyr::mutate(Group = factor(Group))

```

## Random Forest

+ split data
```{r}
set.seed(123)
trainIndex <- caret::createDataPartition(
          MergeData$Group, 
          p = 0.8, 
          list = FALSE, 
          times = 1)

trainData <- MergeData[trainIndex, ]
X_train <- trainData[, -1]
y_train <- trainData[, 1]

testData <- MergeData[-trainIndex, ]
X_test <- testData[, -1]
y_test <- testData[, 1]
```


+ Base Model
```{r}
set.seed(123)
base.fit <- randomForest(
  Group ~ .,
  data = trainData,
  importance = TRUE)

base.fit
```

+ Important score
```{r}
imp_biomarker <- tibble::as_tibble(
    round(importance(base.fit), 2),
    rownames = "FeatureID") %>% 
  dplyr::arrange(desc(MeanDecreaseAccuracy))

# DT::datatable(imp_biomarker)
```

## Boruta
+ feature selection: Boruta Algorithm 

1. First, it duplicates the dataset, and shuffle the values in each column.

2. Then, the algorithm checks for each of your real features if they have higher importance.

3. At every iteration, the algorithm compares the Z-scores of the shuffled copies of the features and the original features to see if the latter performed better than the former (a Z-score is the number of standard deviations from the mean a data point is, for more info click here.).

4. If a feature hasn't been recorded as a hit in say 15 iterations, you reject it and also remove it from the original matrix.


```{r}
set.seed(123)
fs_Boruta <- Boruta::Boruta(
  x = X_train, 
  y = y_train, 
  doTrace = 2)
print(fs_Boruta)

boruta.bank <- Boruta::TentativeRoughFix(fs_Boruta)
print(boruta.bank)

bank_df <- Boruta::attStats(boruta.bank)

plot(boruta.bank, xlab = "", xaxt = "n")
lz <- lapply(1:ncol(boruta.bank$ImpHistory), function(i) {
    boruta.bank$ImpHistory[is.finite(boruta.bank$ImpHistory[, i]), i]
})
names(lz) <- colnames(boruta.bank$ImpHistory)
Labels <- sort(sapply(lz, median))
axis(side = 1, las = 2, labels = names(Labels),
at = 1:ncol(boruta.bank$ImpHistory), cex.axis = 0.7)

# save(X_train, y_train, X_test, y_test, 
#      fs_Boruta, boruta.bank, bank_df,
#      file = paste0(pref, "_RF_Boruta.RData"))

# pdf(paste0(pref_plot, "Fig5_D.pdf"), width = 7, height = 4)
  plot(boruta.bank, xlab = "", xaxt = "n")
  lz <- lapply(1:ncol(boruta.bank$ImpHistory), function(i) {
    boruta.bank$ImpHistory[is.finite(boruta.bank$ImpHistory[, i]), i]
  })
  names(lz) <- colnames(boruta.bank$ImpHistory)
  Labels <- sort(sapply(lz, median))
  axis(side = 1, las = 2, labels = names(Labels),
  at = 1:ncol(boruta.bank$ImpHistory), cex.axis = 0.7)
# dev.off()
```


+ Selected features by Boruta
```{r}
feature_Boruta <- bank_df %>%
  dplyr::filter(decision %in% c("Tentative", "Confirmed"))

trainData_select <- trainData %>%
  dplyr::select(all_of(c("Group", rownames(feature_Boruta))))
X_train_select <- trainData_select[, -1]
y_train_select <- trainData_select[, 1]

testData_select <- testData %>%
  dplyr::select(all_of(c("Group", rownames(feature_Boruta))))
X_test_select <- testData_select[, -1]
y_test_select <- testData_select[, 1]
```


+ Tune parameters to choose the Optimal features:

  - mtry
  
  - ntree
  
```{r}
set.seed(123)

# N-repeat K-fold cross-validation
myControl <- trainControl(
  method = "repeatedcv",
  number = 10,
  repeats = 3,
  search = "random",
  classProbs = TRUE,
  verboseIter = TRUE,
  allowParallel = TRUE)

# customRF
# https://machinelearningmastery.com/tune-machine-learning-algorithms-in-r/
customRF <- list(type = "Classification",
                 library = "randomForest",
                 loop = NULL)

customRF$parameters <- data.frame(
  parameter = c("mtry", "ntree"),
  class = rep("numeric", 2),
  label = c("mtry", "ntree"))

customRF$grid <- function(x, y, len = NULL, search = "grid") {}
customRF$fit <- function(x, y, wts, param, lev, last, weights, classProbs, ...) {
  randomForest(x, y, mtry = param$mtry, ntree=param$ntree, ...)
}

customRF$predict <- function(modelFit, newdata, preProc = NULL, submodels = NULL) {
  predict(modelFit, newdata)
}

customRF$prob <- function(modelFit, newdata, preProc = NULL, submodels = NULL) {
  predict(modelFit, newdata, type = "prob")
}

customRF$sort <- function(x) {x[order(x[, 1]), ]}
customRF$levels <- function(x) {x$classes}

# tuning parameters
tuneGrid <- expand.grid(
  .mtry = c(2:15), # sqrt(ncol(X_train))
  .ntree = seq(500, 2000, 500))


# Register parallel cores
doParallel::registerDoParallel(4)

# train model
set.seed(123)
tune_fit <- train(
  Group ~.,
  data = trainData_select,
  method = customRF, #"rf",
  trControl = myControl,
  tuneGrid = tuneGrid,
  metric = "Accuracy",
  verbose = FALSE) 

## Plot model accuracy vs different values of Cost
print(plot(tune_fit))

## Print the best tuning parameter that maximizes model accuracy
optimalVar <- data.frame(tune_fit$results[which.max(tune_fit$results[, 3]), ])
print(optimalVar)

error.cv <- c()
for (i in 1:5) {
  
  print(i)
  set.seed(i)
  
  fit <- rfcv(trainx = X_train_select, 
              trainy = y_train_select, 
              cv.fold = 5, 
              # scale = "log", 
              step = 0.9,
              ntree = optimalVar$ntree)
  
  error.cv <- cbind(error.cv, fit$error.cv)
}

n.var <- as.numeric(rownames(error.cv))
colnames(error.cv) <- paste("error", 1:5, sep = ".")
err.mean <- apply(error.cv, 1, mean)
err.df <- data.frame(num = n.var, 
                     err.mean = err.mean,
                     error.cv)

# save(tune_fit, file = paste0(pref, "_RF_tuneFit.RData"))
# save(error.cv, file = paste0(pref, "_RF_CVerr.RData"))
```


+ Visualization of Optimal features
```{r, fig.width=6, fig.height=4}
optimal <- err.df$num[which(err.df$err.mean == min(err.df$err.mean))]
print(optimal) ## choose a value
main_theme <- 
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line.x = element_line(linewidth = 0.5, color = "black"),
    axis.line.y = element_line(linewidth = 0.5, color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.title = element_text(size = 11, color = "black", face = "bold"),    
    axis.text = element_text(color = "black", size = 10),
    legend.position = "right",
    legend.background = element_blank(),
    legend.key = element_blank(),
    legend.text = element_text(size = 12),
    text = element_text(family = "serif"))

pl <- 
  ggplot(data = err.df, aes(x = err.df$num)) + 
    geom_line(aes(y = err.df$error.1), color = "grey", linewidth = 0.5) + 
    geom_line(aes(y = err.df$error.2), color = "grey", linewidth = 0.5) +
    geom_line(aes(y = err.df$error.3), color = "grey", linewidth = 0.5) +
    geom_line(aes(y = err.df$error.4), color = "grey", linewidth = 0.5) +
    geom_line(aes(y = err.df$error.5), color = "grey", linewidth = 0.5) +
    geom_line(aes(y = err.df$err.mean), color = "black", linewidth = 0.5) + 
    geom_vline(xintercept = optimal, color = "red", lwd = 0.36, linetype = 2) + 
    coord_trans(x = "log2") +
    scale_x_continuous(breaks = c(1, 5, 10, 20, 30)) +
    labs(x = "Number of Features ", y = "Cross-validation error rate") + 
    annotate("text", 
             x = optimal - 6, 
             y = max(err.df$err.mean), 
             label = paste("Optimal = ", optimal, sep = ""),
             color = "red") +
    main_theme

pl
ggsave(paste0(pref_plot, "optimal_feature_number.pdf"), pl, width = 7, height = 4, dpi = 600)
```


+ Optimal important score
```{r, fig.width=6, fig.height=7}
optimal <- 49

optimal_feature_pl <- imp_biomarker %>%
  dplyr::filter(FeatureID %in% rownames(feature_Boruta)) %>%
  dplyr::slice(1:optimal) %>%
  dplyr::select(FeatureID, MeanDecreaseAccuracy) %>%
  dplyr::arrange(MeanDecreaseAccuracy) %>%
  dplyr::mutate(FeatureID = forcats::fct_inorder(FeatureID)) %>%
  ggplot(aes(x = FeatureID, y = MeanDecreaseAccuracy))+
  geom_bar(stat = "identity", fill = "white", color = "blue") +
  labs(x = "", y = "Mean decrease accuracy") +
  theme(aspect.ratio = 2) +
    coord_flip() +
    main_theme

optimal_feature_pl

ggsave(paste0(pref_plot, "optimal_feature.pdf"), optimal_feature_pl, width = 5, height = 7, dpi = 600)
```



## Final model of selected features

+ building model
```{r}
selected_biomarker <- imp_biomarker %>%
  dplyr::filter(FeatureID %in% rownames(feature_Boruta)) %>%
  dplyr::slice(1:optimal)  

selected_columns <- c("Group", selected_biomarker$FeatureID)

trainData_optimal <- trainData_select %>%
  dplyr::select(all_of(selected_columns))

testData_optimal <- testData_select %>%
  dplyr::select(all_of(selected_columns))

set.seed(123)
rf_fit_optimal <- randomForest(
  Group ~ ., 
  data = trainData_optimal, 
  importance = TRUE, 
  proximity = TRUE,
  ntree = optimalVar$ntree)

rf_fit_optimal
```


+ Confusion matrix

```{r}
group_names <- grp_names
pred_raw <- predict(rf_fit_optimal, newdata = testData_optimal, type = "response")
print(caret::confusionMatrix(pred_raw, testData_optimal$Group))
pred_prob <- predict(rf_fit_optimal, newdata = testData_optimal, type = "prob")  
```


+ ROC

```{r}
AUROC_res <- AUROC(
    DataTest = testData_select, 
    PredProb = pred_prob, 
    label = group_names[1], 
    DataProf = optimal)
AUROC_res$roc_pl
ggsave(paste0(pref_plot, "ROC.pdf"), AUROC_res$roc_pl, width = 5, height = 4, dpi = 600)
save(AUROC_res, file = paste0(pref, "_RF_AUROC.RData"))
```


## Final profile
```{r}
optimal_feature <- imp_biomarker %>%
  dplyr::filter(FeatureID %in% rownames(feature_Boruta)) %>%
  dplyr::slice(1:optimal) %>%
  dplyr::select(FeatureID, MeanDecreaseAccuracy) %>%
  dplyr::arrange(MeanDecreaseAccuracy) %>%
  dplyr::mutate(FeatureID = forcats::fct_inorder(FeatureID))

profile_RF <- profile[pmatch(optimal_feature$FeatureID,
                              rownames(profile)), ]

write.csv(optimal_feature, paste0(pref, "_RF_feature.csv"), row.names = F)
write.table(profile_RF, paste0(pref, "_RF_profile.tsv"), row.names = T, sep = "\t", quote = F)
```

## Session info
```{r}
devtools::session_info()
```


## Reference 

+ [Feature Selection with the Caret R Package](https://machinelearningmastery.com/feature-selection-with-the-caret-r-package/)

+ [Feature selection in r](https://jtr13.github.io/cc21fall2/feature-selection-in-r.html)

+ [Feature Selection in R with the Boruta R Package](https://www.datacamp.com/tutorial/feature-selection-R-boruta)

