---
title: "Machine learning based on Support Vector Machine"
output: html_document
---

## Introduction


Feature selection using Support Vector Machine with Recursive Feature Elimination (SVM_RFE) on the DEGs.


```{r setup, include=TRUE, warning = F, message = F}
knitr::opts_chunk$set(warning = F, message = F)
library(tidyverse)
library(Biobase)
library(ggplot2)
library(data.table)
library(caret)
library(e1071)
library(mlbench)
library(pROC)

library(sysfonts)
library(showtextdb)
library(showtext)
#激活显示中文的包
showtext_auto()

# rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 10000 * 1024^2)
```


# Importing data 
```{r}
grp_names <- c("Healthy", "PAAD")
grp_colors <- c("#1F78B4", "#EE2B2B")

load("../../Result/Exp/exo.RData")
load("../../Result/JiaBo/2DEG/overlap_genes.rda")
source("./utilities.R")
source("../functions.R")
item = "exo" 
profile <- exp_dat_exo
metadata <- phen_dat_exo  
metadata <- metadata %>% dplyr::rename(Group = Compvar) 
pref <- paste0("../../Result/JiaBo/3ML/", item)
fig_plot <- "../../Result/JiaBo/figure/ML1/"
pref_plot <- paste0(fig_plot, item, "_SVM_")

profile_DEG <- profile[pmatch(overlap_genes, rownames(profile)), ] %>%
  t() %>%
  as.data.frame()
```

# prepocess
+ Remove redundant variables

```{r}
feature_cor <- Hmisc::rcorr(as.matrix(profile_DEG), type = "spearman")
index_cor <- caret::findCorrelation(feature_cor$r, cutoff = 0.75, verbose = TRUE)
feature_remain <- feature_cor$r[-index_cor, -index_cor]
profile_remain <- profile_DEG %>% 
  dplyr::select(all_of(rownames(feature_remain))) 
```


+ MergeData 

```{r}
MergeData <- metadata %>%
  dplyr::select(SampleID, Group) %>%
  dplyr::inner_join(profile_remain %>%
                      tibble::rownames_to_column("SampleID"),
                    by = "SampleID") %>%
  tibble::column_to_rownames("SampleID") %>%
  dplyr::mutate(Group = recode(Group, "Healthy" = "Healthy",
                               "PAAD" = "PAAD")) %>%
  dplyr::mutate(Group = factor(Group))
```


## Support Vector Machine

+ split data
```{r}
set.seed(123)
trainIndex <- caret::createDataPartition(
          MergeData$Group, 
          p = 0.8, 
          list = FALSE, 
          times = 1)

trainData <- MergeData[trainIndex, ]
X_train <- trainData[, -1]
y_train <- trainData[, 1]

testData <- MergeData[-trainIndex, ]
X_test <- testData[, -1]
y_test <- testData[, 1]
```


+ Base Model
```{r}
set.seed(123)
base.fit <- e1071::svm(
  Group ~ .,
  data = trainData)

base.fit
```

# RFE
+ feature selection: Recursive Feature Elimination 

1. Rank the importance of all features using the chosen RFE machine learning algorithm.

2. Eliminate the least important feature.

3. Build a model using the remaining features.

4. Repeat steps 1-3 until the desired number of features is reached.



```{r}
set.seed(123)

# y_train必须是factor
y_train <- as.factor(y_train)

# 如果是二分类，建议确认阳性类别在第一个或第二个
table(y_train)

# ===============================
# True SVM-RFE function
# ===============================
svm_rfe <- function(x, y, sizes = c(1:30, 40, 50, 60),
                    cost = 1,
                    nfold = 5,
                    seed = 123) {
  
  set.seed(seed)
  
  x <- as.data.frame(x)
  y <- as.factor(y)
  
  features <- colnames(x)
  ranking <- data.frame(
    FeatureID = character(),
    Rank = integer(),
    Weight = numeric(),
    stringsAsFactors = FALSE
  )
  
  step <- 1
  
  while(length(features) > min(sizes)) {
    
    dat <- data.frame(Class = y, x[, features, drop = FALSE])
    
    fit <- e1071::svm(
      Class ~ .,
      data = dat,
      kernel = "linear",
      cost = cost,
      scale = TRUE
    )
    
    # 计算线性SVM权重：w = t(coefs) %*% SV
    w <- t(fit$coefs) %*% fit$SV
    w <- as.numeric(w)
    names(w) <- features
    
    weight_score <- w^2
    
    # 删除权重最小的一个特征
    remove_feature <- names(which.min(weight_score))
    
    ranking <- rbind(
      ranking,
      data.frame(
        FeatureID = remove_feature,
        Rank = length(features),
        Weight = weight_score[remove_feature]
      )
    )
    
    features <- setdiff(features, remove_feature)
    step <- step + 1
  }
  
  # 剩余特征也加入ranking，权重越大越靠前
  dat <- data.frame(Class = y, x[, features, drop = FALSE])
  fit <- e1071::svm(
    Class ~ .,
    data = dat,
    kernel = "linear",
    cost = cost,
    scale = TRUE
  )
  
  w <- t(fit$coefs) %*% fit$SV
  w <- as.numeric(w)
  names(w) <- features
  weight_score <- sort(w^2, decreasing = FALSE)
  
  remaining_rank <- data.frame(
    FeatureID = names(weight_score),
    Rank = seq(length(features), 1),
    Weight = as.numeric(weight_score)
  )
  
  ranking <- rbind(ranking, remaining_rank)
  
  # 最终排序：Rank越小越重要
  ranking <- ranking %>%
    arrange(Rank)
  
  # ===============================
  # 交叉验证不同特征数量
  # ===============================
  folds <- createFolds(y, k = nfold, returnTrain = FALSE)
  
  cv_results <- data.frame()
  
  for(size in sizes) {
    
    selected_features <- ranking$FeatureID[1:size]
    
    auc_vec <- c()
    acc_vec <- c()
    sens_vec <- c()
    spec_vec <- c()
    
    for(i in seq_along(folds)) {
      
      test_idx <- folds[[i]]
      train_idx <- setdiff(seq_along(y), test_idx)
      
      x_tr <- x[train_idx, selected_features, drop = FALSE]
      y_tr <- y[train_idx]
      x_te <- x[test_idx, selected_features, drop = FALSE]
      y_te <- y[test_idx]
      
      dat_tr <- data.frame(Class = y_tr, x_tr)
      dat_te <- data.frame(x_te)
      
      fit_cv <- e1071::svm(
        Class ~ .,
        data = dat_tr,
        kernel = "linear",
        cost = cost,
        scale = TRUE,
        probability = TRUE
      )
      
      pred_class <- predict(fit_cv, dat_te)
      pred_prob <- attr(predict(fit_cv, dat_te, probability = TRUE), "probabilities")
      
      # 取第二个level作为阳性类，如果你的PDAC不是第二个，请修改这里
      positive_class <- levels(y)[2]
      prob_pos <- pred_prob[, positive_class]
      
      roc_obj <- pROC::roc(
        response = y_te,
        predictor = prob_pos,
        levels = levels(y),
        direction = "<",
        quiet = TRUE
      )
      
      cm <- caret::confusionMatrix(
        pred_class,
        y_te,
        positive = positive_class
      )
      
      auc_vec <- c(auc_vec, as.numeric(pROC::auc(roc_obj)))
      acc_vec <- c(acc_vec, cm$overall["Accuracy"])
      sens_vec <- c(sens_vec, cm$byClass["Sensitivity"])
      spec_vec <- c(spec_vec, cm$byClass["Specificity"])
    }
    
    cv_results <- rbind(
      cv_results,
      data.frame(
        Variables = size,
        AUC = mean(auc_vec, na.rm = TRUE),
        Accuracy = mean(acc_vec, na.rm = TRUE),
        Sensitivity = mean(sens_vec, na.rm = TRUE),
        Specificity = mean(spec_vec, na.rm = TRUE),
        AUCSD = sd(auc_vec, na.rm = TRUE),
        AccuracySD = sd(acc_vec, na.rm = TRUE)
      )
    )
  }
  
  best_size <- cv_results$Variables[which.max(cv_results$AUC)]
  selected_features <- ranking$FeatureID[1:best_size]
  
  final_fit <- e1071::svm(
    Class ~ .,
    data = data.frame(Class = y, x[, selected_features, drop = FALSE]),
    kernel = "linear",
    cost = cost,
    scale = TRUE,
    probability = TRUE
  )
  
  return(list(
    ranking = ranking,
    results = cv_results,
    best_size = best_size,
    selected_features = selected_features,
    final_fit = final_fit
  ))
}

fs_rfe <- svm_rfe(
  x = X_train,
  y = y_train,
  sizes = c(1:30, 40, 50, 60),
  cost = 1,
  nfold = 5,
  seed = 123
)

print(fs_rfe$best_size)

rfe_imp <- fs_rfe$results
p <- ggplot(rfe_imp, aes(x = Variables, y = AUC)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  labs(
    x = "Number of features",
    y = "Cross-validated AUC",
    title = "SVM-RFE feature selection")
print(p)
ggsave(paste0(pref_plot, "optimal_feature.pdf"), p, width = 5, height = 10, dpi = 600)

feature_rfe <- fs_rfe$selected_features
imp_dat <- fs_rfe$ranking %>%
  filter(FeatureID %in% feature_rfe) %>%
  mutate(Importance = 1 / Rank) %>%
  arrange(Rank)
write.csv(imp_dat, paste0(pref, "_SVM_feature.csv"), row.names = F)
```
# Final model of selected features
+ Selected features by rfe
```{r} 
trainData_select <- trainData %>%
  dplyr::select(all_of(c("Group", feature_rfe)))
X_train_select <- trainData_select[, -1]
y_train_select <- trainData_select[, 1]

testData_select <- testData %>%
  dplyr::select(all_of(c("Group", feature_rfe)))
X_test_select <- testData_select[, -1]
y_test_select <- testData_select[, 1]
```


+ Tune parameters:  

  - sigma
  
  - C
  
```{r}
myControl <- trainControl(
  method = "repeatedcv",
  number = 10,
  repeats = 3,
  search = "random",
  summaryFunction = twoClassSummary,  
  classProbs = TRUE,
  verboseIter = TRUE,
  allowParallel = TRUE)

# tuning parameters
tuneGrid <- expand.grid(
  sigma= 2^c(-25, -20, -15,-10, -5, 0), 
  C = 2^c(0:5))

set.seed(123)
tune_fit <- train(
  Group ~.,
  data = trainData_select,
  method = "svmRadial",
  trControl = myControl,
  tuneGrid = tuneGrid)

## Plot model accuracy vs different values of Cost
print(plot(tune_fit))

## Print the best tuning parameter that maximizes model accuracy
optimalVar <- data.frame(tune_fit$results[which.max(tune_fit$results[, 3]), ])
print(optimalVar)

# save(tune_fit, file = paste0(pref, "_SVM_tuneFit.RData"))
```

+ building model
```{r}
set.seed(123)
svm_fit_optimal <- e1071::svm(
  Group ~ ., 
  data = trainData_select, 
  kernel = "radial",
  metric = "ROC",
  gamma = optimalVar$sigma,
  cost = optimalVar$C,
  summaryFunction = twoClassSummary,  
  probability = TRUE)
# svm_fit_optimal
```



+ Confusion matrix

```{r}
group_names <- grp_names
pred_raw <- predict(svm_fit_optimal, newdata = X_test_select, type = "response")
print(caret::confusionMatrix(pred_raw, testData_select$Group))
pred_prob_temp <- predict(svm_fit_optimal, newdata = X_test_select, probability = TRUE)
pred_prob <- attr(pred_prob_temp, "probabilities")
```


+ ROC

```{r}
AUROC_res <- AUROC(
    DataTest = testData_select, 
    PredProb = pred_prob, 
    label = group_names[1], 
    DataProf = fs_rfe$best_size)
AUROC_res$roc_pl
save(AUROC_res, file = paste0(pref, "_SVM_AUROC.RData"))
ggsave(paste0(pref_plot, "ROC.pdf"), AUROC_res$roc_pl, width = 5, height = 4, dpi = 600)
```


# Final profile
```{r}
profile_SVM <- profile[pmatch(feature_rfe, rownames(profile)), ]
write.table(profile_SVM, paste0(pref, "_SVM_profile.tsv"), row.names = T, sep = "\t", quote = F)
```

## Session info
```{r}
devtools::session_info()
```


## Reference 

+ [Feature Selection with the Caret R Package](https://machinelearningmastery.com/feature-selection-with-the-caret-r-package/)

+ [Feature selection in r](https://jtr13.github.io/cc21fall2/feature-selection-in-r.html)

+ [Example for svm feature selection in R](https://stackoverflow.com/questions/17529537/example-for-svm-feature-selection-in-r)

+ [Cannot compute probabilities in SVM?](https://stackoverflow.com/questions/13908267/cannot-compute-probabilities-in-svm)
